---
title: "RNAseq Intuition Plots"
author: "Amanda Charbonneau"
date: '`r Sys.Date()`'
output:
  html_document:
    code_folding: hide
    collapsed: no
    df_print: paged
    number_sections: yes
    theme: cerulean
    toc: yes
    toc_depth: 5
    toc_float: yes
  html_notebook:
    toc: yes
    toc_depth: 5
---

```{r setup, include=FALSE}

library(ggplot2)

library(lme4)

knitr::opts_chunk$set(echo = TRUE, message = F)



```

```{r formatting}

fakeexp <- read.csv("fakeexp.csv")

fakeexp$Treatment <- factor(fakeexp$Treatment, levels = c("black", "green", "blue"))

fakeexp$order <- c(1:144)

fakeexp$center <- scale(fakeexp$order, center = T, scale = F)

fakeexp$Container <- as.factor(fakeexp$Container)

fakeexp$Lane <- as.factor(fakeexp$Lane)

fakeexp$LP_Date <- as.factor(fakeexp$LP_Date)

fakeexp$New_Date <- as.factor(fakeexp$New_Date)

fakeexp$New_Place <- as.factor(fakeexp$New_Place)

fakeexp$Feeding <- as.factor(fakeexp$Feeding)

fakeexp$Prepper <- as.factor(fakeexp$Prepper)

```

## Our experimental data in a  scatterplot

```{r basicscatter}

plot(fakeexp$order, fakeexp$ReadCount, pch=16, xlab="Row Number", ylab = "Read Count")
```

## ReadCount arranged by Row and modeled means

```{r lmScatter}
ggplot(fakeexp, aes(order, ReadCount)) + 
  geom_point() + geom_smooth(method = lm) + 
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black")) + 
  xlab(label = "") + ylab(label = "Read Count")


lm(ReadCount~1, data = fakeexp)

```

## GeneX ~ Feeding and modeled means

```{r ReadbyFeeding}
ggplot(fakeexp, aes(Row, ReadCount, col=Feeding)) + 
  geom_point() + 
    geom_smooth(method = lm) + 
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black")) + 
  xlab(label = "") + ylab(label = "Read Count") 

summary(lm(ReadCount~Feeding, data = fakeexp))


```
## GeneX ~ Treatment and modeled means

```{r ReadbyTreatment}
ggplot(fakeexp, aes(Row, ReadCount, col=Treatment)) + 
  geom_point(color=as.character(fakeexp$color)) + 
    geom_smooth(method = lm) + 
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black")) + 
  xlab(label = "") + ylab(label = "Read Count") +
  scale_color_manual(values = c("black", "darkgreen","blue3"))

summary(lm(ReadCount~Treatment, data = fakeexp))


```

## GeneX ~ Treatment + Feeding and modeled means
```{r ReadbyTreatmentFeeding}
ggplot(fakeexp, aes(Row, ReadCount, col=Treatment, shape=Feeding)) + 
  geom_point(color=as.character(fakeexp$color)) + geom_smooth(method = lm) + 
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black")) + 
  xlab(label = "") + ylab(label = "Read Count") +
  scale_color_manual(values = c("black", "darkgreen","blue3"))

summary(lm(ReadCount~Treatment+Feeding, data = fakeexp))

```

## GeneX ~ Container

```{r ReadbyContainer}
ggplot(fakeexp, aes(Row, ReadCount, col=Container)) + 
  geom_point() + 
    geom_smooth(method = lm) + 
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black")) + 
  xlab(label = "") + ylab(label = "Read Count") 



```
## Dealing with Blocking 

ReadCount~Treatment+Feeding+Container modeled means

```{r RemovingBlockingEffects}
summary(lm(ReadCount~Treatment+Feeding+Container, data = fakeexp))
```

## GeneX ~ Lane and GeneX ~ LP_Date 

With a confounded model of both

```{r ConfoundingBlockingEffects}
ggplot(fakeexp, aes(Row, ReadCount, col=Lane)) + 
  geom_point() + 
    geom_smooth(method = lm) + 
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black")) + 
  xlab(label = "") + ylab(label = "Read Count") 

ggplot(fakeexp, aes(Row, ReadCount, col=LP_Date)) + 
  geom_point() + 
    geom_smooth(method = lm) + 
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black")) + 
  xlab(label = "") + ylab(label = "Read Count") 

summary(lm(ReadCount~Lane+LP_Date, data = fakeexp))


```


## GeneX ~ Feeding and GeneX ~ Prepper


```{r ReadbyTreatmentPrepper}
ggplot(fakeexp, aes(Row, ReadCount, col=Treatment, shape=Prepper)) + 
  geom_point(color=as.character(fakeexp$color)) + geom_smooth(method = lm) + 
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black")) + 
  xlab(label = "") + ylab(label = "Read Count") +
  scale_color_manual(values = c("black", "darkgreen","blue3"))
```


## A confounding-like problem

```{r ImpossibleCombos}
ggplot(fakeexp, aes(Row, ReadCount, col=Lane)) + 
  geom_point() + 
    geom_smooth(method = lm) + 
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black")) + 
  xlab(label = "") + ylab(label = "Read Count") 

ggplot(fakeexp, aes(Row, ReadCount, col=New_Date)) + 
  geom_point() + 
    geom_smooth(method = lm) + 
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black")) + 
  xlab(label = "") + ylab(label = "Read Count") 

summary(lm(ReadCount~Lane+New_Date, data = fakeexp))


```



## GeneX ~ Treatment + Feeding + Container

```{r RunningOutofData}
ggplot(fakeexp, aes(Row, ReadCount, col=Treatment, shape=Feeding)) + 
  geom_point(color=as.character(fakeexp$color)) + geom_smooth(method = lm) + 
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black")) + 
  xlab(label = "") + ylab(label = "Read Count") +
  scale_color_manual(values = c("black", "darkgreen","blue3")) +
 facet_grid(~Container)

summary(lm(ReadCount~Treatment+Feeding+Container, data = fakeexp))

```



## Centered plots to show means more easily

```{r lmScatterModeled}
ggplot(fakeexp, aes(order, ReadCount)) +
  geom_point(color=as.character(fakeexp$color)) + geom_smooth(method = lm) + 
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black")) + 
  xlab(label = "") + ylab(label = "Read Count") +
  geom_vline(xintercept = 72)

lm(ReadCount~1, data = fakeexp)


ggplot(fakeexp, aes(order, ReadCount)) +
  geom_point() + geom_smooth(method = lm) + 
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black")) + 
  xlab(label = "") + ylab(label = "Read Count") +
  geom_vline(xintercept = 72) +
  geom_hline(yintercept = 78.68 )

summary(lm(ReadCount~1, data = fakeexp))

```

```{r ReadbyTreatmentModeled}
ggplot(fakeexp, aes(Row, ReadCount, col=Treatment)) + 
  geom_point(color=as.character(fakeexp$color)) + geom_smooth(method = lm) + 
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black")) + 
  xlab(label = "") + ylab(label = "Read Count") +
  geom_vline(xintercept = 24) +
  geom_vline(xintercept = 72) +
  geom_vline(xintercept = 120)+
   scale_color_manual(values = c("black", "darkgreen","blue3"))


summary(lm(ReadCount~Treatment, data = fakeexp))


```

